Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by the Galaxy Cluster Abell 2744

### 2015-05-19 snTomas.tex
trivial tweak to abstract
small changes in abstract and host mass section

\begin{abstract}
\changenote{Reduced to $<250$ words.}\\
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744
($z=0.308$). This SN 
%has a projected separation from the cluster core
%of $\sim$40\arcsec, 
is closer to the critical lensing line than any
previous cluster-lensed Type Ia SN, 
%As such, it provides 
providing a valuable
opportunity to test galaxy cluster lens models
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  
%We derive a tightly constrained measure
%of the peak apparent magnitude, corrected for light curve shape and
%extinction.  
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of 
$\mu_{\rm obs}=2.03\pm0.29$. 
\change{In future lens models of Abell 2744, this could
provide some leverage for breaking degeneracies that positional
constraints cannot disentangle.}  We compare the measured $\mu$
against predicted magnifications from 17 lens models, \change{ of
which 13 were generated before the SN magnification was known,
qualifying as a pure ``blind test.''}  The models are collectively
fairly accurate: 8 of the models deliver median magnifications that
are consistent with the measured $\mu$ to within 1$\sigma$.  However,
there is a subtle systematic bias: all of the models that disagree
with the SN $\mu$
%by more than $1\sigma$ 
are {\it higher} than the measured value and 6 of these
are $\geq1.5\sigma$ above the measured magnification.  We
evaluate possible causes for this mild bias,
%and reject explanations
%that invoke misinterpretation of the SN data or errors in the
%definition of the lens models. We 
and \changeR{find that model accuracy can be improved to some extent
by using multiply-imaged systems vetted with stringent quality cuts
and requiring a large fraction with spectroscopic redshifts.}
% -- at least for this
%cluster and this line of sight -- 
% A more robust evaluation
%\change{Further examination} of systematic biases 
%could be achieved 
%\change{could be pursued with more}
%lens model comparisons using lensed
%Type Ia SNe, as well as cluster lensing simulations.
\end{abstract}

@@

\change{To measure the stellar mass of the host galaxy
we use Eq. 8 of \citet{Taylor:2011}, which relates the rest-frame
(g-i) color and i-band luminosity to the total stellar mass. To derive
these values, we fixed the redshift at $z=1.3457$ and repeated the SED
fitting using BPZ. From the best-fit SED we extracted rest-frame
optical magnitudes, and corrected them for lensing using a
magnification factor of $\mu=2.0$ -- a value that we will derive from
the SN itself in Section~\ref{sec:DistanceAndMagnification}.  From
this we determine the host galaxy mass to be 
\changeR{$10^{9}$ \Msun.} }

@@

It is now an accepted practice in cosmological analyses using Type Ia
SN to apply a correction to the luminosity of each SN based on the
stellar mass of its host galaxy. Typically this is described as a
simple bifurcation of the SN population: SNe that appear in more
massive hosts are observed to be $\sim$0.08 mag brighter (after
corrections for light curve shape and color) than SNe in low-mass
hosts \citep{Kelly:2010,Sullivan:2010}.  The dividing line for this
purely empirical ``mass step'' correction is generally set around
$10^{10}$\Msun, although this is somewhat arbitrary \citep[see
e.g.][]{Betoule:2014}.

@@

\change{
If we were to apply the correction to this sample, then the SN \tomas\
distance modulus would not be adjusted, because its host galaxy mass
(Section~\ref{sec:HostGalaxy}) is \changeR{below} 10$^{10}$ \Msun. Note that
there is some circularity here, as we used the SN magnification in
deriving the host galaxy mass, and now use the host mass to inform the
magnification.  However, changing the magnification value to $mu=2.5$
\changeR{does not move the \tomas\ host mass above the mass} step threshold, so
the circularity is not relevant in this case. Applying the
mass step correction across the entire control sample would lead to a
small adjustment of the inferred SN \tomas\ magnification, but it is $<5\%$
and therefore could not relieve the observed tension.}

>>>>

\begin{abstract}
\changenote{Reduced to $<250$ words.}\\
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744
($z=0.308$). This SN 
%has a projected separation from the cluster core
%of $\sim$40\arcsec, 
is closer to the critical lensing line than any
previous cluster-lensed Type Ia SN, 
%As such, it provides 
providing a valuable
opportunity to test galaxy cluster lens models
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  
%We derive a tightly constrained measure
%of the peak apparent magnitude, corrected for light curve shape and
%extinction.  
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of 
$\mu_{\rm obs}=2.03\pm0.29$. 
\change{In future lens models of Abell 2744, this could
provide some leverage for breaking degeneracies that positional
constraints cannot disentangle.}  We compare the measured $\mu$
against predicted magnifications from 17 lens models, \change{ of
which 13 were generated before the SN magnification was known,
qualifying as a pure ``blind test.''}  The models are collectively
fairly accurate: 8 of the models deliver median magnifications that
are consistent with the measured $\mu$ to within 1$\sigma$.  However,
there is a subtle systematic bias: all of the models that disagree
with the SN $\mu$
%by more than $1\sigma$ 
are {\it higher} than the measured value and \changeR{8} of these
are $\geq1.5\sigma$ above the measured magnification.  We
evaluate possible causes for this mild bias,
%and reject explanations
%that invoke misinterpretation of the SN data or errors in the
%definition of the lens models. We 
and \changeR{find that model accuracy can be improved to some extent
by using multiply-imaged systems vetted with stringent quality cuts
and requiring that a large fraction have spectroscopic redshifts.}
% -- at least for this
%cluster and this line of sight -- 
% A more robust evaluation
%\change{Further examination} of systematic biases 
%could be achieved 
%\change{could be pursued with more}
%lens model comparisons using lensed
%Type Ia SNe, as well as cluster lensing simulations.
\end{abstract}

@@

\change{To measure the stellar mass of the host galaxy
we use Eq. 8 of \citet{Taylor:2011}, which relates the rest-frame
(g-i) color and i-band luminosity to the total stellar mass. To derive
these values, we fixed the redshift at $z=1.3457$ and repeated the SED
fitting using BPZ. From the best-fit SED we extracted rest-frame
optical magnitudes, and corrected them for lensing using a
magnification factor of $\mu=2.0$ -- a value that we will derive from
the SN itself in Section~\ref{sec:DistanceAndMagnification}.  From
this we determine the host galaxy mass to be 
\changeR{$10^{9.8}$ \Msun.} }

@@

It is now an accepted practice in cosmological analyses using Type Ia
SN to apply a correction to the luminosity of each SN based on the
stellar mass of its host galaxy. Typically this is described as a
simple bifurcation of the SN population: SNe that appear in more
massive hosts are observed to be $\sim$0.08 mag brighter (after
corrections for light curve shape and color) than SNe in low-mass
hosts \citep{Kelly:2010,Sullivan:2010}.  The dividing line for this
purely empirical ``mass step'' correction is generally set around
$10^{10}$\Msun.  \changeR{Although this threshold value} is somewhat arbitrary \citep[see
e.g.][]{Betoule:2014}, \changeR{it happens to be very close to the
-SN \tomas\ host galaxy mass of $10^{9.8}$\Msun (Section~\ref{sec:HostGalaxy}).}

@@

\changeR{
If we were to apply the correction to this sample using the standard
approach, then the SN \tomas\ distance modulus would not be adjusted,
because its host galaxy mass (Section~\ref{sec:HostGalaxy}) is just
below 10$^{10}$ \Msun. Note that there is some circularity here, as we
used the SN magnification in deriving the host galaxy mass, and now
use the host mass to inform the magnification.  However, regardless of
whether the \tomas\ host mass falls above the mass threshold or below,
this correction is not substantial enough to account for the observed
magnification tension.  The appropriate correction to be applied to
a SN above the mass threshold in this redshift range would be $<$0.04
mag \citep{Rigault:2013,Childress:2014b}, which results in a change to
the inferred magnification of \tomas\ that is much less than the
observed discrepancy of $\sim$0.24 mag.}
